Alien species

### 2023-03-23 DRAFT_alien_species.Rmd
even more types and text edits

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(sf)
library(tmap)
library(ggpubr)
library(DT)
library(units)
library(eaTools)
library(tibble)
#library(ggplot2)
#library(tidyr)
#library(dplyr)
knitr::opts_chunk$set(echo = TRUE)

@@

```{r, eval=F}
corrCheck <- st_intersection(alien_data, HIA)
#saveRDS(corrCheck, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/corrCheck_alienSpecies.rds")
saveRDS(corrCheck, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/corrCheck_alienSpecies.rds")
```

@@

And here is the total and relative area of the indicator maps
```{r relative-area, fig.cap="Relative area cover of the indicator maps for the alien species indicator (i.e. value 1 means that all HIAs had > 20 data points, and 0.2 means 20% of them did). If assuming the ecosystems are equally common where they are mapped and where they are not, this measure of data coverage is representative of the real indicator coverage after masking the maps with ecosystem delineations."}
alien_extr %>%
  filter(!is.na(w_mean)) %>%
  mutate(m2 = st_area(.)) %>%
  group_by(eco) %>%
  summarise(sum_area = sum(m2)) %>%
  add_column(total_area = total_area) %>%
  mutate(relative_area = sum_area/total_area) %>%
  ggplot()+
  geom_bar(aes(x = eco, y = relative_area),
           stat = "identity")+
  labs(x = "Ecosystem", y = "Relative indicator data coverage")+
  theme_bw()
```

@@

```{r lat-effect, fig.cap="Effect of latitude on the alien species indicator values. The blue line is a loess smoother (span=0.75)."}
temp2 %>%
  ggplot(aes(x = X, y = w_mean))+
  geom_point()+
  theme_bw()+
  geom_smooth(linewidth=2,
              method="loess",
              span=0.75)+
  labs(x = "Latitude (UTM with offset)",
       y = "Alien species indicator value\n(area weighted)")+
  facet_wrap(~eco)

@@

1. Prepare ecosystem delineation maps in EPSG25833 and perfectly aligned to a master grid

@@

```{r, fig.cap="Indictor values (alien species, extrapolated and area weighted values) in Gran municipality"}
tm_shape(st_downsample(gran_rast, 10))+
  tm_raster(col="w_mean")+
  tm_shape(gran_outline)+
  tm_borders(col="red")
```

Then i mask

@@

```{r, fig.cap="Indictor values (alien species, extrapolated and area weighted values) in Gran municipality"}
tm_shape(st_downsample(gran_rast_masked, 10))+
  tm_raster(col="w_mean",
            palette = "red")+
  tm_shape(gran_outline)+
  tm_borders(col="red")
```

>>>>

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(sf)
library(tmap)
library(ggpubr)
library(DT)
library(units)
library(eaTools)
library(tibble)
library(stars)
knitr::opts_chunk$set(echo = TRUE)

@@

```{r, eval=F}
corrCheck <- st_intersection(alien_data, HIA)
saveRDS(corrCheck, paste0(pData, "cache/corrCheck_alienSpecies.rds")
```

@@

And here is the total and relative area of the indicator maps
```{r relative-area, fig.cap="Relative area cover of the indicator maps for the alien species indicator (i.e. value 1 means that all HIAs had > 20 data points, and 0.2 means 20% of them did). If assuming the ecosystems are equally common where they are mapped and where they are not, this measure of data coverage is representative of the real indicator coverage after masking the maps with ecosystem delineations."}
alien_extr %>%
  filter(!is.na(w_mean)) %>%
  mutate(m2 = st_area(.)) %>%
  group_by(eco) %>%
  summarise(sum_area = sum(m2)) %>%
  add_column(total_area = total_area) %>%
  mutate(relative_area = units::drop_units(sum_area/total_area)) %>%
  ggplot()+
  geom_bar(aes(x = eco, y = relative_area),
           stat = "identity")+
  labs(x = "Ecosystem", y = "Relative indicator data coverage")+
  theme_bw()
```

@@

```{r lat-effect, fig.cap="Effect of latitude on the alien species indicator values. The blue line is a loess smoother (span=0.75).", message=FALSE}
temp2 %>%
  ggplot(aes(x = X, y = w_mean))+
  geom_point()+
  theme_bw()+
  geom_smooth(linewidth=2,
              method="loess",
              span=0.75)+
  labs(x = "Latitude (UTM with offset)",
       y = "Alien species indicator value\n(area weighted)")+
  facet_wrap(~eco)

@@

1. Prepare ecosystem delineation maps in EPSG:25833 and perfectly aligned to a master grid

@@

```{r, fig.cap="Indictor values (alien species, extrapolated and area weighted values) in Gran municipality. Note that the pixelation is very small (10 x 10 m), and that the larger squares are remnants from the HIA map which was 1 x 1 km resolution."}
tm_shape(st_downsample(gran_rast, 10))+
  tm_raster(col="w_mean")+
  tm_shape(gran_outline)+
  tm_borders(col="red")
```

Then i mask to remove areas that are not actually watlands.

@@

```{r, fig.cap="Stratified, area weighted indictor values (alien species) in Gran municipality masked by a wetland ecosystem delimination map."}
tm_shape(st_downsample(gran_rast_masked, 10))+
  tm_raster(col="w_mean",
            palette = "red")+
  tm_shape(gran_outline)+
  tm_borders(col="red")
```
